\documentclass{beamer} % "Beamer" is a word used in Germany to mean video projector. 

\usetheme{metropolis} % Search online for beamer themes to find your favorite or use the Berkeley theme as in this file.


\usepackage{color} % It may be necessary to set PCTeX or whatever program you are using to output a .pdf instead of a .dvi file in order to see color on your screen.
\usepackage{graphicx} % This package is needed if you wish to include external image files.

\theoremstyle{definition} % See Lesson Three of the LaTeX Manual for more on this kind of "proclamation."
\newtheorem*{defn}{Definition} 
\newtheorem{thm}{Theorem}  
\newtheorem{lem}{Lemma}  

\newcommand{\cross}{\otimes{}}

\usepackage{multicol}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{units}
\usepackage{tikz}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\gen{\langle}{\rangle}
\DeclarePairedDelimiter\size{|}{|}
\DeclarePairedDelimiter\abs{|}{|}
\DeclarePairedDelimiter\parens{(}{)}
\setlength{\parindent}{0 pt}
\newcommand{\FF}{\mathcal{F}}
\newcommand{\EE}{\mathcal{E}}
\newcommand{\BB}{\mathcal{B}}
\newcommand{\PP}{\mathcal{P}}
\newcommand{\AAA}{\mathcal{A}}
\newcommand{\II}{\mathcal{I}}


\title{New Static Analysis Techniques to Detect Entropy Failure Vulnerabilities}
\author{Andrew Russell \& Rushi Shah} 
\institute{The University of Texas at Austin}
%\date{} 
% Remove the % from the previous line and change the date if you want a particular date to be displayed; otherwise, today's date is displayed by default.

\AtBeginSection[]  % The commands within the following {} will be executed at the start of each section.
{
\begin{frame} % Within each "frame" there will be one or more "slides."  
\frametitle{Presentation Outline} % This is the title of the outline.
\tableofcontents[currentsection]  % This will display the table of contents and highlight the current section.
\end{frame}
} % Do not include the preceding set of commands if you prefer not to have a recurring outline displayed during your presentation.

\begin{document}


\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Entropy Failures: A Historical Perspective}

	\begin{center}
	\Large{
	\begin{enumerate}
		\item OpenSSL
		\item FreeBSD
	\end{enumerate}
	}
	\end{center}

\end{frame}

\begin{frame}
\frametitle{How Could This Have Been Avoided?}

	- Audit code once, then use relational verification to prove you haven't introduced bugs with small changes to program. 

	- This differential approach works well with the way software is written (CI, etc.)

\end{frame}

\begin{frame}
	\[\text{\Large{Background/Initial Approach}}\]
\end{frame}

\begin{frame}
\frametitle{Taint Analysis}

	- Definition, terminology, uses, etc.

	- Sources are legitimate sources of entropy (/dev/random/)

	- Sinks are things like cryptographic algorithms (KDF)

	- But taint analysis is unsound when used on two versions of the program (overapproximation)

\end{frame}

\begin{frame}
\frametitle{Predicate Abstraction}

	- Finer grained version of taint analysis across versions of a program.

	- Taint set of variable in program two should be a superset of taint set of variable in program one.

	\pause

	- Weaker version: if $v_1$ is the taint set of variable $v$ passed to sink in program one and $v_2$ is the taint set of $v$ in program two, then we would like \[assert(v_1 == v_2)\]

\end{frame}

\begin{frame}
\frametitle{Predicate Abstraction}

	- Off-the-shelf state-of-the-art: CPAChecker

\end{frame}

\begin{frame}
\frametitle{Product Programs}

	- Sequential Composition \[S_1\ ;\ S_2\]

	\pause

	- Synchronized Composition \[S_1 \cross S_2\]

	\pause

	- Hybrid?

\end{frame}

\begin{frame}
	\[\text{\Large{Algorithm}}\]
\end{frame}

\begin{frame}
\frametitle{High level overview}

	\begin{enumerate}
		\item Instrumentation
		\item Product Program
		\item Assertions + CPAChecker
	\end{enumerate}

\end{frame}

\begin{frame}
\frametitle{Instrumentation}
	
	- Replace sources with labelled constants

	- Perform taint analysis on sources to generate environment $\Gamma$ which marks statements involving tainted variables. 

\end{frame}

\begin{frame}
\frametitle{Instrumentation}
	
	- For values that are tainted by more than one source (for example $S_1 + S_2$) replace with one of two uninterpreted functions over the sources: \begin{enumerate}
		\item $preserving(s_1, s_2, \ldots, s_n)$
		\item $nonPreserving(s_1, s_2, \ldots, s_n)$
	\end{enumerate}
	
	- Preserving functions are +, XOR, etc.
	
	- Non-preserving functions are left or right shift, etc.

\end{frame}

\begin{frame}
\frametitle{Sequential Product Program}
	
\end{frame}

\begin{frame}
\frametitle{Naive Synchronized Product Program}
	
\end{frame}

\begin{frame}
\frametitle{Heuristic-Optimized Synchronized Product Program}
	
\end{frame}

\begin{frame}
\frametitle{Assertions + CPAChecker}
	
\end{frame}

\begin{frame}
\frametitle{Correctness}
	
\end{frame}

\begin{frame}
	\[\text{\Large{Future Work}}\]
\end{frame}

\begin{frame}
\frametitle{Implementation Evaluation}
	
\end{frame}

\begin{frame}
	\[\text{\Large{Conclusion}}\]
\end{frame}

\begin{frame}
\frametitle{Acknowledgements}
	\[\text{Prof. Hovav Shacham and Prof. Isil Dillig}\]
\end{frame}

\begin{frame}
\frametitle{References}
	
\end{frame}

\end{document}